How fluid-mechanical erosion creates anisotropic porous media

### 2022-06-27 docs/PhysicsPaper/StokesErosion.tex
edited new sec on tortuosity

\subsection{Tortuosity}

We need to write this section.
Could reference Souzy et al. JFM article on velocity distributions in porous media \cite{souzy2020velocity}

Like the permeability, the tortuosity provides a macroscopic perspective
of the porous medium. We define the tortuosity to be the average length
of streamlines passing through the region $[-1,1] \times [-1,1]$. In
particular, for a particular eroded geometry with velocity field $\uu$,
we calculate streamlines $\pp(t)$ that satisfy
\begin{align}
  \frac{d\pp}{dt} = \uu(\pp), \quad \pp(0) = (-1,y_0),
\end{align}
where $y_0 \in (-1,1)$. Then, the length of a streamline, which only
depends on the initial $y$-coordinate, is
\begin{align}
  \lambda(y_0) = \int_{0}^{t(y_0)} \|\pp(t)\| \,dt.
\end{align}
The integral limit $t(y_0)$ is chosen so that the $x$-coordinate of
$p(t(y_0))$ is $+1$. Then, the longitudinal tortuosity is
\begin{align}
  T_1 = \frac{1}{2}\left(\int_{S}u(-1,y)\lambda(y)\,dy \right)
  \Bigg/
  \left(\int_{S}u(x_0,y)\,dy \right).
  \label{eqn:tortuosity_x}
\end{align} 
The factor of $1/2$ guarantees that $T_1 \geq 1$, and $T_1 = 1$ if and only if no grains are present. Identical to how we calculate the transverse permeability, we rotate each of the rotated bodies by 90 degrees and compute a transverse tortuosity $T_2$. Then, another measure of the anisotropy is
\begin{align}
  \anis = T_2/T_1.
\end{align}
As in the permeability definition of anisotropy, the random configuration of circles used to initialize the erosion simulations will have an anisotropy nearly equal to one. As this configuration erodes, the streamlines in the longitudinal direction are expected to less tortuous than those in the transverse direction, yielding $\anis > 1$.

An alternative and more convenient way to compute the tortuosity is to
use an area integral. If there is no reentrant flow in $D = [-1,1]
\times [-1,1]$, then the longitudinal tortuosity can also be calculated
as~\citep{dud-koz-mat2011}
\begin{align}
  T_1 = \left(\int_D \|\uu(\xx)\|\, d\xx \right) \Bigg/
      \left(\int_D u(\xx)\, d\xx \right).
  \label{eqn:tortuosity_x2}
\end{align}
We note that there may be slow reentrant regions in our geometries~\cite{chiu2020viscous}, but they are sufficiently small that the difference between equations~\eqref{eqn:tortuosity_x} and~\eqref{eqn:tortuosity_x2} can be neglected. We use a similar area integral to compute $T_2$.

>>>>

%-------------------------------------------------------------------%
\subsection{Tortuosity}
%
%We need to write this section.
%Could reference Souzy et al. JFM article on velocity distributions in porous media \cite{souzy2020velocity}

Like the permeability, the tortuosity provides a macroscopic perspective
of the porous medium. We define the tortuosity to be the average length
of streamlines passing through the region $[-1,1] \times [-1,1]$. In
particular, for a particular eroded geometry with velocity field $\uu$,
we calculate streamlines $\pp(t)$ that satisfy
\begin{align}
  \dot{\pp}(t) = \uu(\pp), \quad \pp(0) = (-1,y_0),
\end{align}
where $y_0 \in (-1,1)$. The streamlines are calculated with a fourth-order Runge-Kutta method. Then, the length of each streamline, which only depends on its initial $y$-coordinate, is
\begin{align}
  \lambda(y_0) = \int_{0}^{t(y_0)} \|\pp(t)\| \,dt.
\end{align}
The integral limit $t(y_0)$ is chosen so that the $x$-coordinate of
$\pp(t(y_0))$ is $+1$. Then, the longitudinal tortuosity is
\begin{align}
  T_1 = \frac{1}{2}\left(\int_{S}u(-1,y)\lambda(y)\,dy \right)
  \Bigg/
  \left(\int_{S}u(x_0,y)\,dy \right).
  \label{eqn:tortuosity_x}
\end{align} 
The factor of $1/2$ guarantees that $T_1 \geq 1$, and $T_1 = 1$ if and only if no grains are present. Identical to how we calculate the transverse permeability, we rotate each of the eroded bodies by 90 degrees and compute a transverse tortuosity $T_2$. Then, another measure of the anisotropy is
\begin{align}
  \anis = T_2/T_1.
\end{align}
As in the permeability definition of anisotropy, the random configuration of circles used to initialize the erosion simulations will have an anisotropy nearly equal to one. As this configuration erodes, the streamlines in the longitudinal direction are expected to less tortuous than those in the transverse direction, yielding $\anis > 1$.

An alternative and computationally more convenient way to compute the tortuosity is to use an area integral. If there is no reentrant flow in $D = [-1,1] \times [-1,1]$, then the longitudinal tortuosity can be calculated as~\citep{dud-koz-mat2011}
\begin{align}
  T_1 = \left(\int_D \|\uu(\xx)\|\, d\xx \right) \Bigg/
      \left(\int_D u(\xx)\, d\xx \right).
  \label{eqn:tortuosity_x2}
\end{align}
We use a similar area integral to compute $T_2$. We note that there may be slow reentrant regions in our geometries~\cite{chiu2020viscous}, but they are sufficiently small that the difference between equations~\eqref{eqn:tortuosity_x} and~\eqref{eqn:tortuosity_x2} can be neglected.
